---
title: "Deep Dive: Premiums on Valuation in the Property Partner Portfolio"
output: 
  html_document:
      keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
```

Previously I have looked at visualising the [Property Partner](https://www.propertypartner.co/aboutus#/ourstory) portfolio using [tableau](https://public.tableau.com/views/PropertyPartnerPortfolioVisualisation/PropertyPartnerPortfolio?:embed=y&:display_count=yes), and [explored their resale date from July 2017](https://www.samabbott.co.uk/post/deep-dive-pp-open-house/). In this post I will be exploring the [August Open House resale](https://resources.propertypartner.co/open-house-august-2017/) data focussing on property premiums over both intial and latest valuation. The code for this post is available [here](https://github.com/seabbs/propertypartner/blob/master/report/17-09-18-premium/latest_initial_premium.Rmd).

```{r packages, include = FALSE}
#install.packages("readxl")
library(readxl)

#install.packages("tidyverse")
library(tidyverse)
library(broom)
library(lubridate)
library(stringr)

#install.packages("scales")
library(scales)

#install.packages("plotly")
library(plotly)

#install.packages("ggjoy")
library(ggjoy)

#install.packages("fifer")

#install.packages("viridis")
library(viridis)

#devtools::install_github("dgrtwo/gganimate")
library(gganimate)
```
```{r read-in-the-data, include = FALSE}
#save data from above link into data subfolder of property partner folder
pp_resale <- read_excel("../../data/Purchases-on-Resale-August-2017-Open-House.xlsx")

##Look at data
glimpse(pp_resale)

summary(pp_resale)
```

```{r add-var-resales}
pp_resale <- pp_resale %>% 
  mutate(`Transaction (£)` = `Unit Count` * `Unit Price`) %>% 
  mutate(Year = year(`Last Transaction`), 
         Month = month(`Last Transaction`),
         Week = week(`Last Transaction`), 
         Day = day(`Last Transaction`)) %>% 
  mutate(`Transaction size (£)` = case_when(`Transaction (£)` < 1 ~ "£0-£1",
                                            `Transaction (£)` < 100 ~ "£1-£99",
                                            `Transaction (£)` < 500 ~ "£100-£499",
                                            `Transaction (£)` < 1000 ~ "£500-£999",
                                            `Transaction (£)` < 5000 ~ "£1000-£4,999",
                                            `Transaction (£)` < 50000 ~ "£5,000-£49,999",
                                            `Transaction (£)` >= 50000 ~ "£50,000+") %>% 
           factor(levels = c("£0-£1", "£1-£99", "£100-£499", "£500-£999", "£1000-£4,999", "£5,000-£49,999", "50,000+"))) %>% 
  rowwise() %>% 
  mutate(Area = `Property name` %>% 
           str_split(pattern = ", ") %>% 
           last %>% 
           last) %>% 
  ungroup
```

- Premiums over latest valuation
- Premiums over initial valuation

```{r summarise-premiums}
pp_resale_30_day_premium <- pp_resale %>% 
  mutate(`Transaction Date` = floor_date(`Last Transaction`, unit = "month")) %>% 
  mutate(`Premium to latest share valuation` = `Unit Count` * `Premium to latest share valuation`,
         `Premium to initial share valuation` = `Unit Count` * `Premium to initial share valuation`) %>% 
  group_by(`Transaction Date`, `Property name`, Area) %>% 
  summarise(`Premium to latest share valuation` = sum(`Premium to latest share valuation`)/sum(`Unit Count`),
            `Premium to initial share valuation` = sum(`Premium to initial share valuation`)/sum(`Unit Count`)) %>% 
  group_by(`Property name`) %>% 
  mutate(`Property available for resale` = min(`Transaction Date`))

pp_resale_yearly_premium <- pp_resale %>% 
  mutate(`Transaction Date` = floor_date(`Last Transaction`, unit = "year")) %>% 
  mutate(`Premium to latest share valuation` = `Unit Count` * `Premium to latest share valuation`,
         `Premium to initial share valuation` = `Unit Count` * `Premium to initial share valuation`) %>% 
  group_by(`Transaction Date`, `Property name`, Area) %>% 
  summarise(`Premium to latest share valuation` = sum(`Premium to latest share valuation`)/sum(`Unit Count`),
            `Premium to initial share valuation` = sum(`Premium to initial share valuation`)/sum(`Unit Count`)) %>% 
  group_by(`Property name`) %>% 
  mutate(`Property available for resale` = min(`Transaction Date`))
```

```{r prem-prop-initial}
premium_prop_initial <- pp_resale_30_day_premium %>% 
  ggplot(aes(x = `Transaction Date`, 
             y = `Premium to initial share valuation`,
             colour = `Property name`,
             group = `Property name`)) + 
  geom_point(alpha = 0.8) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.4) + 
  scale_y_continuous(label = percent) + 
  scale_fill_viridis_d() +
  labs(title = "Average Premium over initial Valuation, for each property") +
  theme_minimal() +
  theme(legend.position = "none") 

fig1 <- ggplotly(premium_prop_initial)


fig1

htmlwidgets::saveWidget(fig1, file = "fig1.html", selfcontained = TRUE)
```

```{r prem-prop-latest}
premium_prop_initial <- pp_resale_30_day_premium %>% 
  ggplot(aes(x = `Transaction Date`, 
             y = `Premium to latest share valuation`,
             colour = `Property name`,
             group = `Property name`)) + 
  geom_point(alpha = 0.8) +
  geom_line() +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.4) + 
  scale_y_continuous(label = percent) + 
  scale_fill_viridis_d() +
  labs(title = "Average Premium over latest Valuation, for each property") +
  theme_minimal() +
  theme(legend.position = "none") 

fig2 <- ggplotly(premium_prop_initial)


fig2

htmlwidgets::saveWidget(fig2, file = "fig2.html", selfcontained = TRUE)
```

```{r avg-premium-to-val}
premium_to_share <- pp_resale %>% 
  mutate(`Transaction Date` = floor_date(`Last Transaction`, unit = "month")) %>% 
  mutate(`Premium to latest share valuation` = `Unit Count` * `Premium to latest share valuation`,
         `Premium to initial share valuation` = `Unit Count` * `Premium to initial share valuation`) %>% 
  group_by(`Transaction Date`) %>% 
  summarise(`Premium to latest share valuation` = sum(`Premium to latest share valuation`)/sum(`Unit Count`),
            `Premium to initial share valuation` = sum(`Premium to initial share valuation`)/sum(`Unit Count`)) %>% 
  mutate(`Difference between premium based on initial vs. latest share valuation` = 
           `Premium to initial share valuation` -
           `Premium to latest share valuation`
           ) %>% 
   gather(key = "Share Valuation", value = "Premium",
         `Premium to initial share valuation`,
         `Premium to latest share valuation`)

plot_premium_to_share <- premium_to_share %>% 
  ggplot(aes(x = `Transaction Date`, 
             y = Premium,
             fill = Premium)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(label = percent) + 
  scale_fill_viridis_c() +
  labs(title = "Average Premium over Share Valuation") + 
  theme_minimal() +
  theme(legend.position = "none") + 
  facet_wrap(~`Share Valuation`, ncol = 1)

fig3 <- ggplotly(plot_premium_to_share)


fig3

htmlwidgets::saveWidget(fig3, file = "fig3.html", selfcontained = TRUE)
```

```{r difference-premium-initial-latest}
plot_diff_latest_initial_premium <- premium_to_share %>% 
  ggplot(aes(x = `Transaction Date`, 
             y = `Difference between premium based on initial vs. latest share valuation`,
             fill = `Difference between premium based on initial vs. latest share valuation`)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(label = percent) + 
  scale_fill_viridis_c() +
  labs(title = "Difference in Premiums",
       y = "Difference in Premiums (Initial - Latest)") + 
  theme_minimal() +
  theme(legend.position = "none")

fig4 <- ggplotly(plot_diff_latest_initial_premium)


fig4

htmlwidgets::saveWidget(fig4, file = "fig4.html", selfcontained = TRUE)
```




```{r prem-resale-entry-date}
premium_prop_year_entry <- pp_resale_30_day_premium %>%
  mutate(`Property available for resale` = year(`Property available for resale`) %>% 
           as.character %>% 
           factor) %>% 
  group_by(`Property available for resale`,
           `Transaction Date`) %>% 
  summarise(`Premium to latest share valuation` = 
              mean(`Premium to latest share valuation`),
            `Premium to initial share valuation` = 
              mean(`Premium to initial share valuation`)) %>% 
  gather(key = "Share Valuation", value = "Premium",
         `Premium to initial share valuation`,
         `Premium to latest share valuation`) %>% 
  ggplot(aes(x = `Transaction Date`, 
             y = Premium,
             colour = `Property available for resale`)) + 
  geom_point(alpha = 0.8) +
  geom_line(aes(group = `Property available for resale`)) +
  geom_hline(yintercept = 0, linetype = 2) +
  scale_y_continuous(label = percent) + 
  scale_fill_viridis_d() +
  labs(title = "Average Premium to Latest Valuation by year available for resale") +
  theme_minimal() +
  theme(legend.position = "none") +
    facet_wrap(~`Share Valuation`)

fig5 <- ggplotly(premium_prop_year_entry)

fig5 

htmlwidgets::saveWidget(fig5, file = "fig5.html", selfcontained = TRUE)
```


```{r plot-latest-initial, fig.show = "animate"}
premium_initial_latest <- pp_resale_30_day_premium %>% 
  ggplot(aes(x = `Premium to initial share valuation`, 
             y = `Premium to latest share valuation`,
             colour = `Property name`,
             group = `Property name`,
             frame = `Transaction Date`)) + 
  geom_point(alpha = 0.8) + 
  geom_smooth(aes(group = `Transaction Date`), alpha = 0.2, method = "lm", colour = "#440154CC") +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, linetype = 2, alpha = 0.5) +
  scale_y_continuous(label = percent) + 
  scale_x_continuous(label = percent) + 
  scale_fill_viridis_d() +
  labs(title = "Property Partner Secondary Market",
    subtitle = "Premium on Latest vs. Initial Valuation",
    caption = "Each colour represents a single property in the property partner portfolio. The solid line indicates the linear trend at  a given time.") +
  theme_minimal() +
  theme(legend.position = "none")

gganimate(premium_initial_latest, title_frame = TRUE)
```


```{r plot-latest-initial-year, fig.show = "animate"}
premium_initial_latest_year <- pp_resale_30_day_premium %>%
  mutate(`Property available for resale` = year(`Property available for resale`) %>% 
           factor(ordered = TRUE)) %>% 
  ggplot(aes(x = `Premium to initial share valuation`, 
             y = `Premium to latest share valuation`,
             colour = `Property available for resale`,
             group = `Property available for resale`,
             text = `Property name`,
             frame = `Transaction Date`)) + 
  geom_point(alpha = 0.8) + 
  geom_smooth(aes(group = interaction(`Property available for resale`, `Transaction Date`)), 
              alpha = 0.2, method = "lm", se = FALSE) +
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, 
              linetype = 2, alpha = 0.5) +
  scale_y_continuous(label = percent) + 
  scale_x_continuous(label = percent) + 
  scale_fill_viridis_d() +
  labs(title = "Property Partner Secondary Market",
    subtitle = "Premium on Latest vs. Initial Valuation",
    caption = "Each colour represents a single property in the property partner portfolio. The solid line indicates the linear trend at  a given time.") +
  theme_minimal() +
  theme(legend.position = "bottom")

gganimate(premium_initial_latest_year, title_frame = TRUE)
```

```{r plot-latest-initial-area}
premium_initial_latest_area <- pp_resale_yearly_premium %>%
  mutate(`Transaction Date` = year(`Transaction Date`)) %>%
  ggplot(aes(x = `Premium to initial share valuation`, 
             y = `Premium to latest share valuation`,
             colour = Area,
             text = `Property name`)) + 
  geom_point(alpha = 0.8) + 
  geom_hline(yintercept = 0, linetype = 2, alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = 2, alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, linetype = 2, alpha = 0.5) +
  scale_y_continuous(label = percent) + 
  scale_x_continuous(label = percent) + 
  scale_fill_viridis_d() +
  labs(title = "Property Partner Secondary Market",
    subtitle = "Premium on Latest vs. Initial Valuation",
    caption = "Each colour represents a single property in the property partner portfolio. The solid line indicates the linear trend at  a given time.") +
  theme_minimal() +
  theme(legend.position = "none") +
  facet_wrap(~`Transaction Date`)

fig6 <- ggplotly(premium_initial_latest_area)

fig6

htmlwidgets::saveWidget(fig6, file = "fig6.html", selfcontained = TRUE)
```

```{r premium_by_property, fig.show = "animate", fig.height = 15}
order_properties_by_resale <- pp_resale_30_day_premium %>% 
                    group_by(`Property name`) %>%
                    filter(`Transaction Date` == min(`Transaction Date`)) %>%
                    arrange(desc(`Transaction Date`)) %>% 
                    pull(`Property name`)

premium_property_year <- pp_resale_30_day_premium %>%
   mutate(`Difference (initial - latest)` = 
           `Premium to initial share valuation` -
           `Premium to latest share valuation`
           ) %>%
  ungroup %>% 
  mutate(`Property name` = `Property name` %>% 
           factor(levels = order_properties_by_resale)
  ) %>%
  rename(`Difference (Initial - Latest)` = 
           `Difference (initial - latest)`,
         Initial = `Premium to initial share valuation`,
         Latest = `Premium to latest share valuation`) %>% 
   gather(key = "Share Valuation", value = "Premium",
         Initial,
         Latest,
         `Difference (Initial - Latest)`) %>% 
  mutate(`Property available for resale` = `Property available for resale` %>% 
           year %>% 
           factor(ordered = TRUE)) %>% 
  ggplot(aes(x = `Property name`, 
             y = Premium,
             text = Area,
             fill = `Property available for resale`,
             frame = `Transaction Date`)) + 
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  scale_y_continuous(label = percent) + 
  scale_fill_viridis_d() +
  labs(title = "Premium on Initial vs Latest Valuation") + 
  theme_minimal() +
  coord_flip() +
    theme(legend.position = "bottom",
        axis.text.y = element_text(size = 6) 
        ) +
  facet_wrap(~`Share Valuation`, nrow = 1)

gganimate(premium_property_year, title_frame = TRUE)
```


```{r current-prem-property}
premium_property_current <- pp_resale_30_day_premium %>%
  filter(`Transaction Date` == max(`Transaction Date`)) %>% 
   mutate(`Difference (initial - latest)` = 
           `Premium to initial share valuation` -
           `Premium to latest share valuation`
           ) %>%
  ungroup %>% 
  mutate(`Property name` = `Property name` %>% 
           factor(levels = order_properties_by_resale)
  ) %>%
  rename(`Difference (Initial - Latest)` = 
           `Difference (initial - latest)`,
         Initial = `Premium to initial share valuation`,
         Latest = `Premium to latest share valuation`) %>% 
   gather(key = "Share Valuation", value = "Premium",
         Initial,
         Latest,
         `Difference (Initial - Latest)`) %>% 
  mutate(`Property available for resale` = `Property available for resale` %>% 
           year %>% 
           factor(ordered = TRUE)) %>% 
  ggplot(aes(x = `Property name`, 
             y = Premium,
             text = Area,
             fill = `Property available for resale`)) + 
  geom_bar(stat = "identity", position = "dodge", alpha = 0.8) +
  scale_y_continuous(label = percent) + 
  scale_fill_viridis_d() +
  labs(title = "Current Premium on Initial vs Latest Valuation") +
  coord_flip() +
  theme_minimal() +
    theme(legend.position = "bottom",
        axis.text.y = element_text(size = 6) 
        ) +
  facet_wrap(~`Share Valuation`, nrow = 1)
  
  fig7 <- ggplotly(premium_property_current)

fig7

htmlwidgets::saveWidget(fig7, file = "fig7.html", selfcontained = TRUE)
```